---
title: "DateTime Formats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DateTime Formats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(shiny.CSVImport)
library(knitr)
library(jsonlite)
```

# Date Formats

```{r DateDefinition, include=FALSE}
.DateFormats <- '{
    "af-ZA": "yyyy/MM/dd",
    "am-ET": "d/M/yyyy",
    "ar-AE": "dd/MM/yyyy",
    "ar-BH": "dd/MM/yyyy",
    "ar-DZ": "dd-MM-yyyy",
    "ar-EG": "dd/MM/yyyy",
    "ar-IQ": "dd/MM/yyyy",
    "ar-JO": "dd/MM/yyyy",
    "ar-KW": "dd/MM/yyyy",
    "ar-LB": "dd/MM/yyyy",
    "ar-LY": "dd/MM/yyyy",
    "ar-MA": "dd-MM-yyyy",
    "ar-OM": "dd/MM/yyyy",
    "ar-QA": "dd/MM/yyyy",
    "ar-SA": "dd/MM/yy",
    "ar-SY": "dd/MM/yyyy",
    "ar-TN": "dd-MM-yyyy",
    "ar-YE": "dd/MM/yyyy",
    "arn-CL": "dd-MM-yyyy",
    "as-IN": "dd-MM-yyyy",
    "az-Cyrl-AZ": "dd.MM.yyyy",
    "az-Latn-AZ": "dd.MM.yyyy",
    "ba-RU": "dd.MM.yy",
    "be-BY": "dd.MM.yyyy",
    "bg-BG": "dd.M.yyyy",
    "bn-BD": "dd-MM-yy",
    "bn-IN": "dd-MM-yy",
    "bo-CN": "yyyy/M/d",
    "br-FR": "dd/MM/yyyy",
    "bs-Cyrl-BA": "d.M.yyyy",
    "bs-Latn-BA": "d.M.yyyy",
    "ca-ES": "dd/MM/yyyy",
    "co-FR": "dd/MM/yyyy",
    "cs-CZ": "d.M.yyyy",
    "cy-GB": "dd/MM/yyyy",
    "da-DK": "dd-MM-yyyy",
    "de-AT": "dd.MM.yyyy",
    "de-CH": "dd.MM.yyyy",
    "de-DE": "dd.MM.yyyy",
    "de-LI": "dd.MM.yyyy",
    "de-LU": "dd.MM.yyyy",
    "dsb-DE": "d. M. yyyy",
    "dv-MV": "dd/MM/yy",
    "el-GR": "d/M/yyyy",
    "en-029": "MM/dd/yyyy",
    "en-AU": "d/MM/yyyy",
    "en-BZ": "dd/MM/yyyy",
    "en-CA": "dd/MM/yyyy",
    "en-GB": "dd/MM/yyyy",
    "en-IE": "dd/MM/yyyy",
    "en-IN": "dd-MM-yyyy",
    "en-JM": "dd/MM/yyyy",
    "en-MY": "d/M/yyyy",
    "en-NZ": "d/MM/yyyy",
    "en-PH": "M/d/yyyy",
    "en-SG": "d/M/yyyy",
    "en-TT": "dd/MM/yyyy",
    "en-US": "M/d/yyyy",
    "en-ZA": "yyyy/MM/dd",
    "en-ZW": "M/d/yyyy",
    "es-AR": "dd/MM/yyyy",
    "es-BO": "dd/MM/yyyy",
    "es-CL": "dd-MM-yyyy",
    "es-CO": "dd/MM/yyyy",
    "es-CR": "dd/MM/yyyy",
    "es-DO": "dd/MM/yyyy",
    "es-EC": "dd/MM/yyyy",
    "es-ES": "dd/MM/yyyy",
    "es-GT": "dd/MM/yyyy",
    "es-HN": "dd/MM/yyyy",
    "es-MX": "dd/MM/yyyy",
    "es-NI": "dd/MM/yyyy",
    "es-PA": "MM/dd/yyyy",
    "es-PE": "dd/MM/yyyy",
    "es-PR": "dd/MM/yyyy",
    "es-PY": "dd/MM/yyyy",
    "es-SV": "dd/MM/yyyy",
    "es-US": "M/d/yyyy",
    "es-UY": "dd/MM/yyyy",
    "es-VE": "dd/MM/yyyy",
    "et-EE": "d.MM.yyyy",
    "eu-ES": "yyyy/MM/dd",
    "fa-IR": "MM/dd/yyyy",
    "fi-FI": "d.M.yyyy",
    "fil-PH": "M/d/yyyy",
    "fo-FO": "dd-MM-yyyy",
    "fr-BE": "d/MM/yyyy",
    "fr-CA": "yyyy-MM-dd",
    "fr-CH": "dd.MM.yyyy",
    "fr-FR": "dd/MM/yyyy",
    "fr-LU": "dd/MM/yyyy",
    "fr-MC": "dd/MM/yyyy",
    "fy-NL": "d-M-yyyy",
    "ga-IE": "dd/MM/yyyy",
    "gd-GB": "dd/MM/yyyy",
    "gl-ES": "dd/MM/yy",
    "gsw-FR": "dd/MM/yyyy",
    "gu-IN": "dd-MM-yy",
    "ha-Latn-NG": "d/M/yyyy",
    "he-IL": "dd/MM/yyyy",
    "hi-IN": "dd-MM-yyyy",
    "hr-BA": "d.M.yyyy.",
    "hr-HR": "d.M.yyyy",
    "hsb-DE": "d. M. yyyy",
    "hu-HU": "yyyy. MM. dd.",
    "hy-AM": "dd.MM.yyyy",
    "id-ID": "dd/MM/yyyy",
    "ig-NG": "d/M/yyyy",
    "ii-CN": "yyyy/M/d",
    "is-IS": "d.M.yyyy",
    "it-CH": "dd.MM.yyyy",
    "it-IT": "dd/MM/yyyy",
    "iu-Cans-CA": "d/M/yyyy",
    "iu-Latn-CA": "d/MM/yyyy",
    "ja-JP": "yyyy/MM/dd",
    "ka-GE": "dd.MM.yyyy",
    "kk-KZ": "dd.MM.yyyy",
    "kl-GL": "dd-MM-yyyy",
    "km-KH": "yyyy-MM-dd",
    "kn-IN": "dd-MM-yy",
    "ko-KR": "yyyy-MM-dd",
    "kok-IN": "dd-MM-yyyy",
    "ky-KG": "dd.MM.yy",
    "lb-LU": "dd/MM/yyyy",
    "lo-LA": "dd/MM/yyyy",
    "lt-LT": "yyyy.MM.dd",
    "lv-LV": "yyyy.MM.dd.",
    "mi-NZ": "dd/MM/yyyy",
    "mk-MK": "dd.MM.yyyy",
    "ml-IN": "dd-MM-yy",
    "mn-MN": "yy.MM.dd",
    "mn-Mong-CN": "yyyy/M/d",
    "moh-CA": "M/d/yyyy",
    "mr-IN": "dd-MM-yyyy",
    "ms-BN": "dd/MM/yyyy",
    "ms-MY": "dd/MM/yyyy",
    "mt-MT": "dd/MM/yyyy",
    "nb-NO": "dd.MM.yyyy",
    "ne-NP": "M/d/yyyy",
    "nl-BE": "d/MM/yyyy",
    "nl-NL": "d-M-yyyy",
    "nn-NO": "dd.MM.yyyy",
    "nso-ZA": "yyyy/MM/dd",
    "oc-FR": "dd/MM/yyyy",
    "or-IN": "dd-MM-yy",
    "pa-IN": "dd-MM-yy",
    "pl-PL": "yyyy-MM-dd",
    "prs-AF": "dd/MM/yy",
    "ps-AF": "dd/MM/yy",
    "pt-BR": "d/M/yyyy",
    "pt-PT": "dd-MM-yyyy",
    "qut-GT": "dd/MM/yyyy",
    "quz-BO": "dd/MM/yyyy",
    "quz-EC": "dd/MM/yyyy",
    "quz-PE": "dd/MM/yyyy",
    "rm-CH": "dd/MM/yyyy",
    "ro-RO": "dd.MM.yyyy",
    "ru-RU": "dd.MM.yyyy",
    "rw-RW": "M/d/yyyy",
    "sa-IN": "dd-MM-yyyy",
    "sah-RU": "MM.dd.yyyy",
    "se-FI": "d.M.yyyy",
    "se-NO": "dd.MM.yyyy",
    "se-SE": "yyyy-MM-dd",
    "si-LK": "yyyy-MM-dd",
    "sk-SK": "d. M. yyyy",
    "sl-SI": "d.M.yyyy",
    "sma-NO": "dd.MM.yyyy",
    "sma-SE": "yyyy-MM-dd",
    "smj-NO": "dd.MM.yyyy",
    "smj-SE": "yyyy-MM-dd",
    "smn-FI": "d.M.yyyy",
    "sms-FI": "d.M.yyyy",
    "sq-AL": "yyyy-MM-dd",
    "sr-Cyrl-BA": "d.M.yyyy",
    "sr-Cyrl-CS": "d.M.yyyy",
    "sr-Cyrl-ME": "d.M.yyyy",
    "sr-Cyrl-RS": "d.M.yyyy",
    "sr-Latn-BA": "d.M.yyyy",
    "sr-Latn-CS": "d.M.yyyy",
    "sr-Latn-ME": "d.M.yyyy",
    "sr-Latn-RS": "d.M.yyyy",
    "sv-FI": "d.M.yyyy",
    "sv-SE": "yyyy-MM-dd",
    "sw-KE": "M/d/yyyy",
    "syr-SY": "dd/MM/yyyy",
    "ta-IN": "dd-MM-yyyy",
    "te-IN": "dd-MM-yy",
    "tg-Cyrl-TJ": "dd.MM.yy",
    "th-TH": "d/M/yyyy",
    "tk-TM": "dd.MM.yy",
    "tn-ZA": "yyyy/MM/dd",
    "tr-TR": "dd.MM.yyyy",
    "tt-RU": "dd.MM.yyyy",
    "tzm-Latn-DZ": "dd-MM-yyyy",
    "ug-CN": "yyyy-M-d",
    "uk-UA": "dd.MM.yyyy",
    "ur-PK": "dd/MM/yyyy",
    "uz-Cyrl-UZ": "dd.MM.yyyy",
    "uz-Latn-UZ": "dd/MM yyyy",
    "vi-VN": "dd/MM/yyyy",
    "wo-SN": "dd/MM/yyyy",
    "xh-ZA": "yyyy/MM/dd",
    "yo-NG": "d/M/yyyy",
    "zh-CN": "yyyy/M/d",
    "zh-HK": "d/M/yyyy",
    "zh-MO": "d/M/yyyy",
    "zh-SG": "d/M/yyyy",
    "zh-TW": "yyyy/M/d",
    "zu-ZA": "yyyy/MM/dd"
  }'

DateFormats <- jsonlite::fromJSON(.DateFormats)
```

The list if date formats has been taken from StackOverflow. It looks like this have `r length(DateFormats)` entries in total. The number of unique formats is `r length(unique(unlist(DateFormats)))`.

```{r}
kable(head(as.data.frame(unlist(DateFormats))))
```

The unique date formats are:

```{r echo=FALSE}
Frequencies <- table(unlist(DateFormats))
Frequencies <- as.data.frame(Frequencies)
Frequencies <- Frequencies[order(Frequencies[[2]], decreasing = TRUE), ]
colnames(Frequencies) <- c("Format", "Frequency")
kable(Frequencies)
```

Based on these data I chose the following formats as default. All formats with 10 occurrences or more are selected without question. I chose a few more that seemed pretty much redundant given the list which was already there. The standard format specified by ISO 8601 (yyyy-mm-dd) is among the top 10, too.

```{r echo=FALSE}
SelectedDates <- Frequencies[c(1:6), ]
SelectedDates <- rbind(SelectedDates, Frequencies[Frequencies[["Format"]] == "dd/MM/yy", ])
SelectedDates <- rbind(SelectedDates, Frequencies[Frequencies[["Format"]] == "dd.MM.yy", ])
SelectedDates[["R-Format"]] <- SelectedDates[["Format"]]
SelectedDates[["R-Format"]] <- gsub("d{1,2}", "%d", SelectedDates[["R-Format"]])
SelectedDates[["R-Format"]] <- gsub("MM", "%m", SelectedDates[["R-Format"]])
SelectedDates[["R-Format"]] <- gsub("yyyy", "%Y", SelectedDates[["R-Format"]])
SelectedDates[["R-Format"]] <- gsub("yy", "%y", SelectedDates[["R-Format"]])

DateCoverage <- sum(SelectedDates[["Frequency"]]) / sum(Frequencies[["Frequency"]])

kable(SelectedDates)
```

The total coverage of default formats is `r format(DateCoverage, digits = 2)`

```{r}
SelectedDates <- SelectedDates[ order(SelectedDates[["R-Format"]]), ]
Choices <- SelectedDates[["R-Format"]]
names(Choices) <- SelectedDates[["Format"]]
dput(Choices)
```

# Time

Time formats seem simpler. Standard ISO format is HH:MM:SS. In most English-speaking regions no 24 hour notation is used.

```{r}

```

# Formatting Strings

For a better usability it makes more sense to use a ["dd.mm.yyyy" format syntax](https://docs.microsoft.com/en-us/dotnet/standard/base-types/custom-date-and-time-format-strings) instead of `R`s POSIX notation.

| Format       | POSIX Format | Description             | Examples                        |
|:-----------  |:-----------  |:------------------------|:--------------------------------|
| d            | %e           | Day of the month, 1-31  | 2009-06-01T13:45:30 -\> 1  <br> 2009-06-15T13:45:30 -\> 15 |
| dd           | %d           | Day of the month, 01-31 | 2009-06-01T13:45:30 -\> 01 <br> 2009-06-15T13:45:30 -\> 15 |
| ddd          | %a           | Abbreviated name of the day of the week | 2009-06-15T13:45:30 -\> Mon (en-US)  <br> 2009-06-15T13:45:30 -\> Пн (ru-RU) <br> 2009-06-15T13:45:30 -\> lun. (fr-FR) |
| dddd         | %A           | Full name of the day of the week        | 2009-06-15T13:45:30 -\> Monday (en-US)  <br>  2009-06-15T13:45:30 -\> понедельник (ru-RU)  <br>  2009-06-15T13:45:30 -\> lundi (fr-FR) |
|              | %u           | Weekday as a decimal number (1--7, Monday is 1) |                             |
|              | %w           | Weekday as decimal number (0--6, Sunday is 0)   |                             |
| f            | -            | Tenths of a second                       | 2009-06-15T13:45:30.6170000 -\> 6 <br> 2009-06-15T13:45:30.05 -\> 0 |
| ff           | -            | Hundredths of a second                   | 2009-06-15T13:45:30.6170000 -\> 61 <br> 2009-06-15T13:45:30.0050000 -\> 00 |
| fff          | -            | Milliseconds                             | 6/15/2009 13:45:30.617 -\> 617 <br> 6/15/2009 13:45:30.0005 -\> 000    |
| ffff         | -            | Ten thousandths of a second              | 2009-06-15T13:45:30.6175000 -\> 6175 <br> 2009-06-15T13:45:30.0000500 -\> 0000 |
| ffffff       | -            | Millionths of a second                   | 2009-06-15T13:45:30.6175425 -\> 6175425 <br> 2009-06-15T13:45:30.0001150 -\> 0001150 |
| F            | -            | Tenths of a second (if non-zero)         | 2009-06-15T13:45:30.6170000 -\> 6 <br> 2009-06-15T13:45:30.0500000 -\> (no output) |
| FF           | -            | Hundredths of a second (if non-zero)     | 2009-06-15T13:45:30.6170000 -\> 61 <br> 2009-06-15T13:45:30.0050000 -\> (no output) |
| FFF          | -            | Milliseconds (if non-zero)               | 009-06-15T13:45:30.6170000 -\> 617 <br> 2009-06-15T13:45:30.0005000 -\> (no output) |
| FFFF         | -            | Ten thousandths of a second (if non-zero) | 2009-06-15T13:45:30.5275000 -\> 5275 <br> 2009-06-15T13:45:30.0000500 -\> (no output) |
| FFFFF        | -            | Hundred thousandths of a second (if non-zero) | 2009-06-15T13:45:30.6175400 -\> 61754 <br> 2009-06-15T13:45:30.0000050 -\> (no output) |
| FFFFFF       | -            | Millionths of a second (if non-zero)   | 2009-06-15T13:45:30.6175420 -\> 617542 <br> 2009-06-15T13:45:30.0000005 -\> (no output) |
| FFFFFFF      | -            | Ten millionths of a second (if non-zero) | 2009-06-15T13:45:30.6175425 -\> 6175425 <br> 2009-06-15T13:45:30.0001150 -\> 000115 |
| ss.f[f[f[f]]]| %OSn         | Seconds and fractions. `n` specifies 10E-n | |
| g, gg        |              | Period or era                        | 2009-06-15T13:45:30.6170000 -\> A.D. |
| h            | %l \*        | Hour, on a 12-hour clock 1-12        | 2009-06-15T01:45:30 -\> 1  |
| hh           | %I           | Hour, on a 12-hour clock 01-12       | 2009-06-15T01:45:30 -\> 01 |
| H            | %k \*        | Hour, on a 24-hour clock 0-23        | 2009-06-15T01:45:30 -\> 1 <br> 2009-06-15T13:45:30 -\> 13 |
| HH           | %H           | Hour, on a 24-hour clock 00-23       | 2009-06-15T01:45:30 -\> 01 <br> 2009-06-15T13:45:30 -\> 13 |
| K            |              | Time zone information                   | 2009-06-15T13:45:30, Kind Unspecified -\>  <br>  2009-06-15T13:45:30, Kind Utc -\> Z  <br>  2009-06-15T13:45:30, Kind Local -\> -07:00 (depends on local computer settings)  <br>  With DateTimeOffset values:  <br>  2009-06-15T01:45:30-07:00 --\> -07:00  <br>  2009-06-15T08:45:30+00:00 --\> +00:00 |
| m            |              | Minute, 0-59                            | 2009-06-15T13:29:30 -\> 29 |
| mm           | %M           | Minute, 00-59                           | 2009-06-15T01:45:30 -\> 45 |
| M            |              | Month, 1-12                             | 2009-06-15T13:45:30 -\> 6  |
| MM           | %m           | Month, 01-12                            | 2009-06-15T13:45:30 -\> 06 |
| MMM          | %b           | Abbreviated name of month               | 2009-06-15T13:45:30 -\> Jun (en-US)  <br>  2009-06-15T13:45:30 -\> juin (fr-FR)  <br>  2009-06-15T13:45:30 -\> Jun (zu-ZA)  |
| MMMM         | %B           | Full name of month                      | 2009-06-15T13:45:30 -\> June (en-US)  <br>  2009-06-15T13:45:30 -\> juni (da-DK)  <br>  2009-06-15T13:45:30 -\> uJuni (zu-ZA)|
| s            | %S (input only)| Second, 0-59                          | 2012-06-15T13:45:09 -\> 9  |
| ss           | %S           | Second, 00-59                           | 2012-06-15T13:45:09 -\> 09 |
| t            |              | First character of the AM/PM designator | 2009-06-15T13:45:30 -\> P (en-US)  <br>  2009-06-15T13:45:30 -\> 午 (ja-JP)  <br>  2009-06-15T13:45:30 -\> (fr-FR)    |
| tt           | %p           | AM/PM designator                        | 2009-06-15T13:45:30 -\> PM (en-US)  <br>  2009-06-15T13:45:30 -\> 午後 (ja-JP)  <br>  2009-06-15T13:45:30 -\> (fr-FR)    |
| y            | %y (input only) | Year, 0-99                           | 0001-01-01T00:00:00 -\> 1 <br>  0900-01-01T00:00:00 -\> 0   <br>  1900-01-01T00:00:00 -\> 0   <br>  2009-06-15T13:45:30 -\> 9 <br>  2019-06-15T13:45:30 -\> 19 |
| yy           | %y           | Year, 00-99                             | 0001-01-01T00:00:00 -\> 01  <br>  0900-01-01T00:00:00 -\> 00  <br>  1900-01-01T00:00:00 -\> 00  <br>  2019-06-15T13:45:30 -\> 19    |
| yyy          | n/a          | Year, with a minimum of three digits    | 0001-01-01T00:00:00 -\> 001 <br> 0900-01-01T00:00:00 -\> 900 <br> 1900-01-01T00:00:00 -\> 1900 <br> 2009-06-15T13:45:30 -\> 2009  |
| yyyy         | %Y           | Year as a four-digit number             | 0900-01-01T00:00:00 -\> 0900  <br>  1900-01-01T00:00:00 -\> 1900  <br>  2009-06-15T13:45:30 -\> 2009  |
| yyyyy        |              | Year as a five-digit number             | 0001-01-01T00:00:00 -\> 00001  <br>  2009-06-15T13:45:30 -\> 02009 |
| z            |              | Hours offset from UTC, without leading zeros | 2009-06-15T13:45:30-07:00 -\> -7 |
| zz           |              | Hours offset from UTC, with leading zero for a single-digit value | 009-06-15T13:45:30-07:00 -\> -07 |
| zzz          |              | Hours and minutes offset from UTC      | 2009-06-15T13:45:30-07:00 -\> -07:00 |
|              | %z           | Signed offset in hours and minutes from UTC, so -0800 is 8 hours behind UTC. | |
| :            |              | Time separator                         | 2009-06-15T13:45:30 -\> : (en-US) |
|              |              |                                        | 2009-06-15T13:45:30 -\> . (it-IT) |
|              |              |                                        | 2009-06-15T13:45:30 -\> : (ja-JP) |
| /            |              | Date separator                         | 2009-06-15T13:45:30 -\> / (en-US) |
|              |              |                                        | 2009-06-15T13:45:30 -\> - (ar-DZ) |
|              |              |                                        | 2009-06-15T13:45:30 -\> . (tr-TR) |
| "string"     |              | Literal string delimiter               | 2009-06-15T13:45:30 ("arr:" h:m t) -\> arr: 1:45 P |
|              |              |                                        | |
| 'string'     |              |                                        | 2009-06-15T13:45:30 ('arr:' h:m t) -\> arr: 1:45 P |
| \%           |              | Defines the following character as a custom format specifier | 2009-06-15T13:45:30 (%h) -\> 1 |
| \\           |              | Escape character                       | 2009-06-15T13:45:30 (h \h) -\> 1 h |
| Any other character |       | The character is copied to the result string unchanged | 2009-06-15T01:45:30 (arr hh:mm t) -\> arr 01:45 A |
\* Not in the POSIX standards and less widely implemented



### POSIX Combination Formats

| Format      | POSIX Format | Description                            | Examples                        |
|-------------|:-------------|:---------------------------------------|:--------------------------------|
| \-          | %c           | Date and time. Locale-specific on output, "%a %b %e %H:%M:%S %Y" on input. | |
| \-          | %D           | Date format such as %m/%d/%y: the C99 standard says it should be that exact format (but not all OSes comply). | |
| yyyy-mm-dd  | %F           | %Y-%m-%d  | |
| hh:mm:ss tt | %r           | For output, the 12-hour clock time (using the locale's AM or PM): only defined in some locales, and on some OSes misleading in locales which do not define an AM/PM indicator. For input, equivalent to %I:%M:%S %p. | |
| HH:mm       | %R           | %H:%M     | |
| HH:mm:ss    | %T           | %H:%M:%S  | |
| yy/mm/dd    | %x           | Date.     | |
|             |              | Locale-specific on output, "%y/%m/%d" on input. |   |
| (HH:mm:ss)  | %X           | Time. Locale-specific on output, "%H:%M:%S" on input. |  |
| \-          | %+           | (Output only) Similar to %c, often "%a %b %e %H:%M:%S %Z %Y". May depend on locale. | |


### POSIX only

| Format      | POSIX Format | Description                            | Examples                        |
|-------------|:-------------|:---------------------------------------|:--------------------------------|
|             | %g            | last two digits of the week-based year (see %V)    | |
|             | %G            | week-based year (see %V) as a decimal number       | |
|             | %j            | Day of year as decimal number (001--366): For input, 366 is only valid in a leap year.| |
| \-          | %n            | Newline on output, arbitrary white space on input. | |
|             | %t            | Tab on output, arbitrary white space on input.     | |
|             | %U            | Week of the year as decimal number (00--53) using Sunday as the first day 1 of the week | |
|             | %W            | Week of the year as decimal number (00--53) using Monday as the first day of week       | |
|             | %s \*         |                                                                                         | |

\* Not in the POSIX standards and less widely implemented

